
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(rnassqs)
library(maps)
library(mapdata)
```

# Weather Data Analysis

### Load the PRISM daily maximum temperatures

```{r tmax data}

# daily max temperature
# dimensions: counties x days x years
prism <- readMat("data/prismiowa.mat")

# look at county #1
t_1981_c1 <- prism$tmaxdaily.iowa[1,,1]
t_1981_c1[366]
plot(1:366, t_1981_c1, type = "l")

ggplot() +
  geom_line(mapping = aes(x=1:366, y = t_1981_c1)) +
  theme_bw() +
  xlab("day of year") +
  ylab("daily maximum temperature (째C)") +
  ggtitle("Daily Maximum Temperature, Iowa County #1")


```


```{r tidying up}

# assign dimension names to tmax matrix
dimnames(prism$tmaxdaily.iowa) <- list(prism$COUNTYFP, 1:366, prism$years)

# converted 3d matrix into a data frame
tmaxdf <- as.data.frame.table(prism$tmaxdaily.iowa)

# relabel the columns
colnames(tmaxdf) <- c("countyfp","doy","year","tmax")
tmaxdf <- tibble(tmaxdf)

```

## Temperature trends

### Summer temperature trends: Winneshiek County

```{r temp trends, message=FALSE}

tmaxdf$doy <- as.numeric(tmaxdf$doy)
tmaxdf$year <- as.numeric(as.character(tmaxdf$year))

winnesummer <- tmaxdf %>%
  filter(countyfp==191 & doy >= 152 & doy <= 243) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnesummer, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (째C)") +
  geom_smooth(method = lm)

lm_summertmax <- lm(meantmax ~ year, winnesummer)
summary(lm_summertmax)

```

### Winter Temperatures - Winneshiek County

```{r winter temps, message=F}

winnewinter <- tmaxdf %>%
  filter(countyfp==191 & (doy <= 59 | doy >= 335) & !is.na(tmax)) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnewinter, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (째C)") +
  geom_smooth(method = lm)

lm_wintertmax <- lm(meantmax ~ year, winnewinter)
summary(lm_wintertmax)

```

### Multiple regression -- Quadratic time trend

```{r quadratic temp trend}

winnewinter$yearsq <- winnewinter$year^2

lm_wintertmaxquad <- lm(meantmax ~ year + yearsq, winnewinter)
summary(lm_wintertmaxquad)
winnewinter$fitted <- lm_wintertmaxquad$fitted.values

ggplot(winnewinter) +
  geom_point(mapping = aes(x = year, y = meantmax)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "tmax")

```

### Download NASS corn yield data

```{r yield download, results='hide'}

# set our API key with NASS
nassqs_auth(key = "DD0322AC-2B3C-34A0-B93C-52D5A8844518")

# parameters to query on 
params <- list(commodity_desc = "CORN", util_practice_desc = "GRAIN", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
cornyieldsall <- nassqs_yields(params)

cornyieldsall$county_ansi <- as.numeric(cornyieldsall$county_ansi)
cornyieldsall$yield <- as.numeric(cornyieldsall$Value)

# clean and filter this dataset
cornyields <- select(cornyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
cornyields <- tibble(cornyields)

```

## Assignment

### Question 1a: Extract Winneshiek County corn yields, fit a linear time trend, make a plot. Is there a significant time trend?

```{r, message=FALSE}
# county data
cornWinneshiek<-cornyields%>%
  filter(county_name == "WINNESHIEK")

# linear model
lm_Winneshiek <- lm(yield ~ year, cornWinneshiek)
summary(lm_Winneshiek)

# plot
ggplot(cornWinneshiek, mapping = aes(x = year, y = yield)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(x = "Year", y = "Yield")

```

\hstart

* There is a significant increase in yield over time.

\hstop

### Question 1b: Fit a quadratic time trend (i.e., year + year^2) and make a plot. Is there evidence for slowing yield growth? 

```{r}
#year^2
cornWinneshiek$yearsq <- cornWinneshiek$year^2

# linear model
lm_Winneshiek <- lm(yield ~ year + yearsq, cornWinneshiek)
summary(lm_Winneshiek)
cornWinneshiek$fitted <- lm_Winneshiek$fitted.values


# plot
ggplot(cornWinneshiek) +
  geom_point(mapping = aes(x = year, y = yield)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "Year", y = "Yield")
```

\hstart

* The quadratic time trend does not show evidence of slowing yield growth.

\hstop

### Question 2 -- Time Series: Let's analyze the relationship between temperature and yields for the Winneshiek County time series. Use data on yield and summer avg Tmax. Is adding year or Tmax^2 to your model helpful? Make a plot and interpret the results.

```{r}
# join data
WinneshiekJoin<-winnesummer%>%
  inner_join(cornWinneshiek, by = "year")


# linear model
WinneshiekJoin$tempsq <-WinneshiekJoin$meantmax^2
lm_Winne <- lm(yield ~ meantmax + tempsq, WinneshiekJoin)
summary(lm_Winne)


# plot
WinneshiekJoin$quadfitted<-lm_Winne$fitted.values
ggplot(WinneshiekJoin) +
  geom_point(mapping = aes(x = meantmax, y = yield)) +
  geom_line(mapping = aes(x = meantmax, y = quadfitted)) +
  theme_bw() +
  labs(x = "Temperature", y = "Yield")

```

\hstart

* It is not helpful to add Tmax^2 to the model. 

\hstop

### Question 3 -- Cross-Section: Analyze the relationship between temperature and yield across all counties in 2018. Is there a relationship? Interpret the results.

```{r}
# get 2018 data
cyield18 <- cornyields %>% 
  filter(year == 2018) %>% 
  group_by(county_name) %>% 
  unique() %>% 
  filter(!is.na(county_ansi))

temp18 <- tmaxdf%>%
  group_by(countyfp)%>%
  filter(year == 2018)%>%
  filter(!is.na(tmax))%>%
  filter(doy >= 152 & doy <= 243) %>%
  summarize(meantmax = mean(tmax)) %>%
  rename(county_ansi = "countyfp")

temp18$county_ansi <- as.numeric(as.character(temp18$county_ansi))

#Joining the two dfs together
county_yield18 <- left_join(cyield18,temp18, by='county_ansi')

# lm
county_yield18$tempsq <-county_yield18$meantmax^2
lm_county18 <- lm(yield ~ meantmax + tempsq, county_yield18)
summary(lm_county18)

county_yield18$quadfitted<-lm_county18$fitted.values

#plot
ggplot(county_yield18, aes(x = meantmax, y = yield))+
  geom_point()+
  theme_bw()+
  labs(x = "Temperature", y = "Yield")

#plot quadratic
ggplot(county_yield18) +
  geom_point(mapping = aes(x = meantmax, y = yield)) +
  geom_line(mapping = aes(x = meantmax, y = quadfitted)) +
  theme_bw() +
  labs(x = "Temperature", y = "Yield")

```

\hstart

* There is not a trend but the quadratic, however, does show a trend. Looking at the quadratic, there is a decrease in yield for both cooler and hotter temperatures, the peak yield occurs closer to 28째C.

\hstop

### Question 4 -- Panel: One way to leverage multiple time series is to group all data into what is called a "panel" regression. Convert the county ID code ("countyfp" or "county_ansi") into factor using as.factor, then include this variable in a regression using all counties' yield and summer temperature data. How does the significance of your temperature coefficients (Tmax, Tmax^2) change? Make a plot comparing actual and fitted yields and interpret the results of your model.

```{r, message=FALSE}
# combine data
ttocombine <- tmaxdf%>%
  filter(doy >= 152 & doy <= 243) %>%
  group_by(countyfp, year)%>%
  rename(county_ansi = countyfp)%>%
  summarize(meantmax = mean(tmax))

#factor
ttocombine$county_ansi <- as.numeric(as.character(ttocombine$county_ansi))

#Joining the data
yieldtemp <- inner_join(cornyields,ttocombine,)%>%
  unique()

#square temp
yieldtemp$meantmaxsq <- yieldtemp$meantmax^2

#lm
lm_yieldtemp<-lm(yield ~county_ansi + meantmax + meantmaxsq + year, yieldtemp)
summary(lm_yieldtemp)

#fitted values
yieldtemp$fittedyield<-lm_yieldtemp$fitted.values

#plot
ggplot(yieldtemp, aes(x = fittedyield, y = yield))+
  geom_point()+
  geom_abline(color = "purple", size = 1)+
  labs(x = "Fitted Yield Values", y = "Yield")

```

\hstart

* Because the fitted values generally follow the actual yield values and there is a high p-value there is a significant goodness of fit for the model.

\hstop

### Question 5 -- Soybeans: Download NASS data on soybean yields and explore either a time series relationship for a given county, the cross-sectional relationship for a given year, or a panel across all counties and years.

```{r, results='hide', message=FALSE}

# set our API key with NASS
nassqs_auth(key = "DD0322AC-2B3C-34A0-B93C-52D5A8844518")

# parameters to query on 
params2 <- list(commodity_desc = "SOYBEANS", statisticcat_desc = "YIELD", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
soyyieldsall <- nassqs_yields(params)

soyyieldsall$county_ansi <- as.numeric(soyyieldsall$county_ansi)
soyyieldsall$yield <- as.numeric(soyyieldsall$Value)

# clean and filter this dataset
soyyields <- select(soyyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
soyyields <- tibble(soyyields)

# county data
soyWinneshiek<-soyyields%>%
  filter(county_name == "WINNESHIEK")

# linear model
lm_soy <- lm(yield ~ year, soyWinneshiek)
summary(lm_soy)

# plot
ggplot(soyWinneshiek, mapping = aes(x = year, y = yield)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(x = "Year", y = "Yield")

```



### Bonus: Find a package to make a county map of Iowa displaying some sort of information about yields or weather. Interpret your map.


### Bonus #2: Challenge question - map trends in corn yields by county across Iowa. Interpret your map.
